Early Observations of the Type Ia Supernova \abc:\\ XXX NEED SOMETHING CLEVER TO SAY HERE XXX

### 2017-08-08 doc/ms.tex
Final proof read edits to make sure new draft flows from beginning to end.

Early observations of Type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia, \abc. The initial detection was made only $0.15\pm^{0.07}_{0.15}$~d after SN first light, the time when radiation from $^{56}$Ni decay has diffused to the SN photosphere. In the $\sim$24 hr after discovery, \abc\ rose by $\sim$2 mag, following a near-linear rise in flux for $\sim$4 d. 
% Assuming the photospheric velocity follows a $t^{-0.22}$ power law, the time of explosion for \abc\ is approximately equal to the time of first light.
Strong \ion{C}{2} absorption is detected in the early spectra of \abc, before disappearing after $\sim$7~d. Unlike SN\,2011fe, the $(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after explosion. We show that the fast, near-linear rise in flux for \abc\ cannot be explained by either SN shock breakout and the associated, subsequent cooling or the collision of the SN ejecta with a stellar companion. Instead, we argue that the early characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors, and (iii) the strong lines from ionized carbon in the earliest spectra, are the result of either vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or the interaction of the ejecta with diffuse material, or a combination of the two. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.

@@

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitor systems and explosion
physics are still debated (see a recent review by
\citealt{2014ARA&A..52..107M}). Extremely detailed
observations in the hours to days after explosion are one of the most promising avenues to further
constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a star with $R_\ast \lesssim 0.02\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass, \citeauthor{2012ApJ...744L..17B} derive the mean density of the progenitor star, confirming that at least some Type Ia SNe come from compact and degenerate stars.

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate binary
may detect excess emission, relative to most Type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a
SN ejecta-companion collision (but see also \citealt{2016MNRAS.459.4428K,2017arXiv170603613N}). While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$ 
10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
(\citealt{2016ApJ...820...92M}; though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction), and SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$ light curves \citep{2017arXiv170608990H}. 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

@@

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark period after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark period is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta. In the case of strong mixing, the dark period is short, or 
non-existent, as $\gamma$-rays from the $^{56}$Ni decay diffuse to the photosphere rapidly. This also leads to blue optical colors and a rapid initial rise in the observed light curve. If the mixing is weak and the $^{56}$Ni is confined to the innermost layers of the ejecta, the dark period can last for several days as the radioactive energy diffuses to the photosphere. The early evolution of such SNe results in redder colors and a more moderate rise in luminosity \citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic Type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.\footnote{While examing a sample of Type Ib/c SNe, \citet{2015A&A...574A..60T} found that a significant dark phase is not prevalent for core-collapse SNe.}

@@

During the spring of 2016, the iPTF survey observed the field of \abc\ every night in either the $g_\mathrm{PTF}$- or 
$R_\mathrm{PTF}$-band.\footnote{P48 observations of \abc\ are reported in the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters throughout, which are similar to the SDSS $g'$ and $r'$ filters, respectively (see \citealt{2012PASP..124..854O} for details on PTF calibration). The correction from the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters to SDSS $g'$ and $r'$ requires knowledge of the intrinsic source color (see Eqns.~1 and 2 in \citealt{2012PASP..124..854O}). The spectral diversity of SNe Ia in the days after explosion is poorly constrained, and as a result the color terms for \abc\ at these epochs are unknown. We proceed by assuming the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ calibration is on the AB system, which strictly speaking is incorrect, but this does not fundamentally alter any of our conclusions.} 
Survey observations were conducted with the
CFH12K camera \citep{2008SPIE.7014E..4YR} on the Palomar Observatory 
48-inch telescope (P48; \citealt{2009PASP..121.1395L}). Images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN 
\citep{2017PASP..129a4002M}. The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

@@

The Reionization and Transients InfraRed (RATIR) camera on the autonomous 1.5\,m Harold L. Johnson Telescope at the Observatorio Astron\'{o}mico Nacional \citep{2012SPIE.8446E..10B,2012SPIE.8444E..5LW} was used to observe \abc\ in the $r'i'ZYJH$ filters. Given the lack of a cold shutter in RATIR’s design, IR dark frames are not available. Laboratory testing, however, confirms that the dark current is negligible in both IR detectors \citep{2012SPIE.8453E..1OF}.

The RATIR data were reduced, coadded, and analysed using standard CCD and IR processing techniques in \texttt{IDL}, \texttt{Python}, \texttt{SExtractor} \citep{1996A&AS..117..393B}, and \texttt{SWarp}. Aperture photometry is obtained following the methods described in \citet{2014AJ....148....2L}. The $r'i'Z$ filters are calibrated to SDSS \citep{2012ApJS..203...21A}, while the $JH$ filters are calibrated to the Two Micron All Sky Survey \citep{2006AJ....131.1163S}. For the $Y$-band calibration, we used an empirical relation in terms of the $J$ and $H$ magnitudes derived from the United Kingdom Infrared Telescope (UKIRT; \citealt{2007A&A...467..777C}) Wide Field Camera observations \citep{2009MNRAS.394..675H}.

The \textit{Swift} satellite observed \abc\ on 14 epochs, beginning 
$\sim$15 d pre-maximum light through $\sim$22 d post maximum. The SN 
flux is measured via aperture photometry on Ultraviolet-Optical
Telescope (UVOT) images via the usual procedures in 
\texttt{HEASoft}, including corrections for the coincident loss and 
aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. There are no pre-SN UVOT images at the 
SN location in the \textit{Swift} archive.  Visual inspection of the
UVOT images suggests negligible host-galaxy contamination in our 
UVOT flux measurements. No X-ray emission is detected from \abc\ by the \textit{Swift} X-ray Telescope (XRT).

@@

Spectroscopic observations of \abc\ were taken with a variety
of telescopes and instruments over multiple epochs spanning from a
couple of days after explosion to two months after $B$-band maximum. 
An observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}/\texttt{Matlab}. The optical
spectral evolution of \abc\ is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

@@

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.8\textwidth]{spectra.pdf}
  \caption{Observed spectral sequence of \abc. The spectra are 
  normalized by their median flux between 6,000 and 
  7,000$\,\textrm{\AA}$.  The phase of each spectrum 
  relative to $T_{B_\mathrm{max}}$ is shown. Telluric absorption
  bands are grayed out. Line identifications are provided for 
  the spectral features discussed in the text. 
  For clarity, high-resolution spectra obtained with the VLT have been omitted (see Ferretti et al.\ 2017, submitted, for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

@@

A detailed study of the reddening towards \abc\ is presented in a 
companion paper (Ferretti et al.\ 2017, submitted). Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in Ferretti et al.\ (2017) show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying a large amount of extinction (e.g., \citealt{2012MNRAS.426.1465P}), Ferretti et al.\ compare the evolution of \abc\ to the well-observed normal Type Ia SN\,2011fe and find evidence for only a small amount of extinction. The empirical relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and \citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a poor tracer of reddening in Type Ia SNe. Thus, we instead adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ (Ferretti et al.\ 2017). For the remainder of our analysis we therefore adopt a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$.

@@

To determine the brightness and time of $B$-band maximum for 
\abc, we fit the P60 light curves with the \texttt{sncosmo} software package.\footnote{The
  \texttt{sncosmo} Python module is available at
  \url{https://sncosmo.readthedocs.io/en/v1.5.x/}.} This fit includes a \texttt{SALT2} template \citep{2007A&A...466...11G} that has been corrected for extinction using the \citep{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V) = 0.08 \; \mathrm{mag}$.

We determine the time of rest-frame \textit{B}-band maximum to be 
 $\textrm{MJD}_{max}=57499.54\pm0.23$, the coefficient
of the zeroth principle component $x_0 = 0.0086 \pm 0.0003$, the
coefficient of the first principle component $x_1 = 0.96 \pm 0.15$, 
and the color term $c = 0.033 \pm 0.029$. The best-fit model also 
gives an unreddened apparent peak magnitude of $m^*_{B}=15.80 \pm 
0.04 \,\textrm{mag}$ in the SN rest frame.

In the following sections, we define the best-fit
value $\textrm{MJD}_{max}=57499.54$ as the time of $B$-band maximum, $T_{B_\mathrm{max}}$, which we also adopt as phase $t=0$.

After establishing \abc\ as a normal SN Ia, we use the latest
calibration \citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} using $m^*_{B}$, $x_1$ and $c$ to derive a distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ to the SN, provided that the host galaxy of \abc\ has a stellar mass less than $10^{10}\sm$. We note that a more massive host galaxy would result in a larger inferred distance modulus that is nevertheless consistent within the uncertainties. For the following analysis we adopt a distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ for \abc.

@@

The time of first light for SNe is usually estimated by 
extrapolating early-phase light curves to determine when the 
SN flux is equal to 0. Assuming an ideal expanding fireball with 
constant temperature, \citet{1982ApJ...253..785A} derives that $f 
\propto t^2$, where $f$ is the SN flux and $t$ is the time since 
explosion. Despite the basic assumption of a constant temperature at 
early times, multiple studies have found that the early emission 
from Type Ia SNe can be described as a power law in time, with 
power-law index consistent with 2, i.e. $f \propto t^2$ (e.g., \citealt{2006AJ....132.1707C, 2010ApJ...712..350H, 2011MNRAS.416.2607G}).\footnote{Many of the studies that find $f \propto t^2$ sample SNe Ia light curves at a phase that is closer to $T_{B,\mathrm{max}}$ than the initial observations of \abc\ discussed here.}

As our observations include especially early observations of \abc, there are upper limits $\sim$1 d prior to the discovery epoch, we model the early flux from \abc\ as a power law, but allow the power-law index to vary, as opposed to fixing it at 2, to account for potential variations in the photospheric temperature during expansion:
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t \le t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ ,
\end{equation}
where $t_0$ is the time of first light, $\alpha$ is the power-law index, and $t$ is measured in the rest-frame of the SN. To determine $t_0$ and $\alpha$ we fit the earliest observations of \abc. Due to slight variations in the passbands, we fit the model only to the relative flux measured in the $g_\mathrm{PTF}$-band, which is the only filter with observations prior to first light, a necessity for constraining $t_0$.

To determine the best fit parameters, we search a large grid over $t_0$, $\alpha$
and the proportionality constant, and minimize $\chi^2$. 
The modeling results show that the SN flux rises approximately
linearly between
$t=-18\,\textrm{d}$ and $t=-14\,\textrm{d}$. Figure \ref{fig:early_lc_fit} shows the best-fit result and
the joint marginal distribution of $t_0$ and $\alpha$. From the
best-fit model we obtain $\alpha=0.98 \pm ^{0.16}_{0.14}$ and $t_0=-17.91 \pm ^{0.07}_{0.15}$ d, where the uncertainties represent the marginalized 95\% confidence intervals. Our first
detection of \abc\ occurred $\sim$${0.15}\,\textrm{d}$ after 
the SN first light. 
% Figure~\ref{fig:early_lc_fit} additionally shows $g'$-bands observations from P60 and LCO, where we have normalized the flux assuming that the $g$ magnitudes from each instrument have the same zero-point. This assumption is incorrect (see \S\ref{sec:obs}), and, as expected, the residuals show a systematic offset between the $g_\mathrm{PTF}$-band and the $g'$-band. Nevertheless, if we ignore this systematic and fit all of the $g$-band observations simultaneously we obtain marginalized best-fit parameters that are consistent within the uncertainties with the $g_\mathrm{PTF}$-only values above.
In the analysis that follows, the precise values of the best-fit parameters are not important. The critical finding here is that $\alpha \approx 1$ and $t_0 \approx -18 \, \mathrm{d}$.

As previously noted, a precise determination of the rise time, $t_\mathrm{rise}$, of SNe Ia is challenging as there may be a ``dark phase'' following explosion \citep{2014ApJ...784...85P}. Nevertheless, to be consistent with previous studies (e.g., \citealt{2011MNRAS.416.2607G}), here we find $t_rise = 17.91\pm _{0.07}^{0.15}$ d based on our fit for $t_0$. We caution that this measurement may be underestimated due to the uncertain time of explosion.

>>>>

Early observations of Type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia, \abc. By fitting a power law to our early light curve, we infer that first light for the SN, that is when appreciable heating from $^{56}$Ni decay has first diffused to the photosphere, only occurred $0.15\pm_{0.07}^{0.15}$~d before our first detection. In the $\sim$24 hr after discovery, \abc\ rose by $\sim$2 mag, following a near-linear rise in flux for $\sim$4 d. 
% Assuming the photospheric velocity follows a $t^{-0.22}$ power law, the time of explosion for \abc\ is approximately equal to the time of first light.
Strong \ion{C}{2} absorption is detected in the early spectra of \abc, before disappearing after $\sim$7~d. Unlike SN\,2011fe, the $(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after explosion. We show that our early observations of \abc\ cannot be explained by either SN shock breakout and the associated, subsequent cooling, or the SN ejecta colliding with a stellar companion. Instead, we argue that the early characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors, and (iii) the strong absorption from ionized carbon, are the result of either vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or ejecta interaction with diffuse material, or a combination of the two. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.

@@

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitor systems and explosion
physics are still debated (see a recent review by
\citealt{2014ARA&A..52..107M}). Extremely detailed
observations in the hours to days after explosion provide a promising avenue to further constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a star with $R_\ast \lesssim 0.02\;\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass, \citeauthor{2012ApJ...744L..17B} derive the mean density of the progenitor star, confirming that at least some Type Ia SNe come from compact and degenerate stars.

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate binary
may detect excess emission, relative to most Type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a SN ejecta-companion collision (but see also \citealt{2016MNRAS.459.4428K,2017arXiv170603613N}). While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
(\citealt{2016ApJ...820...92M}; though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction), and SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$ light curves \citep{2017arXiv170608990H}. 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

@@

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark phase after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark phase is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta. In the case of strong mixing, the dark phase is short, or 
non-existent, as $\gamma$-rays from the radioactive $^{56}$Ni diffuse to the photosphere rapidly. This also leads to blue optical colors and a rapid initial rise in the observed light curve. If the mixing is weak and the $^{56}$Ni is confined to the innermost layers of the ejecta, the dark phase can last for several days as the radioactive energy diffuses to the photosphere. The early evolution of such SNe results in redder colors and a more moderate rise in luminosity \citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic Type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.\footnote{\citet{2015A&A...574A..60T} found that a significant dark phase is not prevalent for a sample of Type Ib/c SNe.}

@@

During the spring of 2016, the iPTF survey observed the field of \abc\ every night during dark time in either the $g_\mathrm{PTF}$ or 
$R_\mathrm{PTF}$ band.\footnote{P48 observations of \abc\ are reported in the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters throughout, which are similar to the SDSS $g'$ and Mould-$R$ filters, respectively (see \citealt{2012PASP..124..854O} for details on PTF calibration). The correction from the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters to SDSS $g'$ and $r'$ requires knowledge of the intrinsic source color (see Eqns.~1 and 2 in \citealt{2012PASP..124..854O}). The spectral diversity of SNe Ia in the days after explosion is poorly constrained, and as a result the color terms for \abc\ at these epochs are unknown. We proceed by assuming the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ calibration is on the AB system, which strictly speaking is incorrect, but this does not fundamentally alter any of our conclusions.} 
Survey observations were conducted with the
CFH12K camera \citep{2008SPIE.7014E..4YR} on the Palomar Observatory 
48-inch telescope (P48; \citealt{2009PASP..121.1395L}). Images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN 
\citep{2017PASP..129a4002M}. The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

@@

The Reionization and Transients InfraRed (RATIR) camera on the autonomous 1.5\,m Harold L. Johnson Telescope at the Observatorio Astron\'{o}mico Nacional \citep{2012SPIE.8446E..10B,2012SPIE.8444E..5LW} was used to observe \abc\ in the $r'i'ZYJH$ filters. By design, RATIR lacks a cold shutter, which means IR dark frames are not available. Laboratory testing, however, confirms that the dark current is negligible in both IR detectors \citep{2012SPIE.8453E..1OF}.

The RATIR data were reduced, coadded, and analyzed using standard CCD and IR processing techniques in \texttt{IDL}, \texttt{Python}, \texttt{SExtractor} \citep{1996A&AS..117..393B}, and \texttt{SWarp}. Aperture photometry is obtained following the methods described in \citet{2014AJ....148....2L}. The $r'i'Z$ filters are calibrated to SDSS \citep{2014ApJS..211...17A}, while the $JH$ filters are calibrated to the Two Micron All Sky Survey \citep{2006AJ....131.1163S}. For the $Y$-band calibration, we used an empirical relation in terms of the $J$ and $H$ magnitudes derived from the United Kingdom Infrared Telescope (UKIRT; \citealt{2007A&A...467..777C}) Wide Field Camera observations \citep{2009MNRAS.394..675H}.

The \textit{Swift} satellite observed \abc\ on 14 epochs, beginning 
$\sim$15 d pre-maximum light through $\sim$22 d post maximum. The SN 
flux is measured via aperture photometry on Ultraviolet-Optical
Telescope (UVOT) images via the usual procedures in 
\texttt{HEASoft}, including corrections for coincidence loss and 
aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. There are no pre-SN UVOT images at the 
SN location in the \textit{Swift} archive.  Visual inspection of the
UVOT images suggests negligible host-galaxy contamination in our 
UVOT flux measurements. No X-ray emission is detected from \abc\ by the \textit{Swift} X-ray Telescope (XRT).

@@

Spectroscopic observations of \abc\ were taken with a variety
of telescopes and instruments over multiple epochs spanning from a
couple of days after discovery to two months after $B$-band maximum. 
An observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}/\texttt{Matlab}. The optical
spectral evolution of \abc\ is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

@@

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.78\textwidth]{spectra.pdf}
  \caption{Observed spectral sequence of \abc. The spectra are 
  normalized by their median flux between 6,000 and 
  7,000$\,\textrm{\AA}$.  The phase of each spectrum 
  relative to $T_{B_\mathrm{max}}$ is shown. Telluric absorption
  bands are grayed out. Line identifications are provided for 
  the spectral features discussed in the text. 
  For clarity, high-resolution spectra obtained with the VLT have been omitted (see Ferretti et al.\ 2017, submitted, for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

@@

A detailed study of the reddening towards \abc\ is presented in a 
companion paper (Ferretti et al.\ 2017, submitted). Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in Ferretti et al.\ (2017) show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying a large amount of extinction (e.g., \citealt{2012MNRAS.426.1465P}), Ferretti et al.\ compare the evolution of \abc\ to the well-observed normal Type Ia SN\,2011fe and find evidence for only a small amount of extinction. The empirical relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and \citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a poor tracer of reddening in Type Ia SNe. Thus, we instead adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ (Ferretti et al.\ 2017). Thus, we assume a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$ for the remainder of our analysis.

@@

To determine the brightness and time of $B$-band maximum for 
\abc, we fit the P60 light curves with the \texttt{sncosmo} software package.\footnote{\texttt{sncosmo} is available at
  \url{https://sncosmo.readthedocs.io}.} This fit includes a \texttt{SALT2} template \citep{2007A&A...466...11G} that has been corrected for extinction using the \citep{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V) = 0.08 \; \mathrm{mag}$.

We determine the time of rest-frame \textit{B}-band maximum to be 
 $\textrm{MJD}_\mathrm{max}=57499.54\pm0.23$, the coefficient
of the zeroth principle component $x_0 = 0.0086 \pm 0.0003$, the
coefficient of the first principle component $x_1 = 0.96 \pm 0.15$, 
and the color term $c = 0.033 \pm 0.029$. The best-fit model also 
gives an unreddened apparent peak magnitude of $m^*_{B}=15.80 \pm 
0.04 \,\textrm{mag}$ in the SN rest frame. In the following sections, we define the best-fit value $\textrm{MJD}_\mathrm{max}=57499.54$ as the time of $B$-band maximum, $T_{B_\mathrm{max}}$, which we also define as phase $t=0$.

We measure the (pseudo)bolometric luminosity, $L_{UVOIR}$, of \abc\ at peak via trapezoidal integration of the reddening-corrected flux from the UV, optical, and NIR (UVOIR) filters. The light curves in the individual filters are interpolated so that $L_{UVOIR}$ is evaluated at common epochs in each filter. From this integration, we determine that \abc\ reaches a maximum luminosity $L_\mathrm{max} = 1.2 \pm 0.1 \times 10^{43} \; \mathrm{erg \; s}^{-1}$. This value is consistent with the normal SNe Ia studied in  \citet{2016A&A...588A..84D}, a sample which excludes low-luminosity SNe.

After establishing \abc\ as a normal SN Ia, we use the latest
calibration \citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} using $m^*_{B}$, $x_1$ and $c$ to derive a distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ to the SN, provided that the host galaxy of \abc\ has a stellar mass $< 10^{10}\sm$. A more massive host galaxy would result in a larger inferred distance modulus that is nevertheless consistent within the uncertainties. For the following analysis we adopt a distance modulus $\mu = 34.89 \pm 0.10 \,\textrm{mag}$ for \abc.

@@

The time of first light, defined as the moment when appreciable heating from $^{56}$Ni decay has first diffused to the photosphere, for SNe is usually estimated by 
extrapolating early-phase light curves to determine when the 
SN flux is equal to 0. Assuming an ideal, expanding fireball with 
constant temperature, \citet{1982ApJ...253..785A} derives that $f 
\propto t^2$, where $f$ is the SN flux and $t$ is the time since 
explosion. Despite the basic assumption of a constant temperature at 
early times, multiple studies have found that the early emission 
from Type Ia SNe can be described as a power law in time, with 
power-law index consistent with 2, i.e. $f \propto t^2$ (e.g., \citealt{2006AJ....132.1707C, 2010ApJ...712..350H, 2011MNRAS.416.2607G}).\footnote{Many of the studies that find $f \propto t^2$ sample SNe Ia light curves at a phase that is closer to $T_{B,\mathrm{max}}$ than the initial observations of \abc\ discussed here.}

As our observations include especially early observations of \abc, there are upper limits $\sim$1 d prior to the discovery epoch, we model the early flux from \abc\ as a power law, but allow the power-law index to vary, as opposed to fixing it at 2, to account for potential variations in the photospheric temperature during expansion:
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t \le t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ ,
\end{equation}
where $t_0$ is the time of first light, $\alpha$ is the power-law index, and $t$ is measured in the SN rest-frame. To determine $t_0$ and $\alpha$ we fit the earliest observations of \abc. Due to slight variations in the passbands, we fit the model only to the relative flux measured in the $g_\mathrm{PTF}$-band, which is the only filter with observations prior to first light, a necessity for constraining $t_0$.

To determine the best fit parameters, we search a large grid over $t_0$, $\alpha$,
and the proportionality constant, and minimize $\chi^2$. 
The modeling results show that the SN flux rises approximately
linearly between
$t=-18\,\textrm{d}$ and $t=-15\,\textrm{d}$. Figure \ref{fig:early_lc_fit} shows the best-fit result and
the joint marginal distribution of $t_0$ and $\alpha$. From the
best-fit model we obtain $\alpha=0.98 \pm ^{0.16}_{0.14}$ and $t_0=-17.91 \pm ^{0.07}_{0.15}$ d, where the uncertainties represent the marginalized 95\% confidence intervals. Our first
detection of \abc\ occurred $\sim$${0.15}\,\textrm{d}$ after 
the SN first light. 
% Figure~\ref{fig:early_lc_fit} additionally shows $g'$-bands observations from P60 and LCO, where we have normalized the flux assuming that the $g$ magnitudes from each instrument have the same zero-point. This assumption is incorrect (see \S\ref{sec:obs}), and, as expected, the residuals show a systematic offset between the $g_\mathrm{PTF}$-band and the $g'$-band. Nevertheless, if we ignore this systematic and fit all of the $g$-band observations simultaneously we obtain marginalized best-fit parameters that are consistent within the uncertainties with the $g_\mathrm{PTF}$-only values above.
In the analysis that follows, the precise values of the best-fit parameters are not important. The critical finding here is that $\alpha \approx 1$ and $t_0 \approx -18 \, \mathrm{d}$.

As previously noted, a precise determination of the rise time, $t_\mathrm{rise}$, of SNe Ia is challenging as there may be a ``dark phase'' following explosion \citep{2014ApJ...784...85P}. Nevertheless, to be consistent with previous studies (e.g., \citealt{2011MNRAS.416.2607G}), here we find $t_\mathrm{rise} = 17.91\pm _{0.07}^{0.15}$ d based on our fit for $t_0$. We caution that this measurement may be underestimated due to the uncertain time of explosion.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.48\textwidth]{early_lc.pdf}
  \caption{Best-fit power-law model to describe the early flux from 
  \abc\ in the $g_\mathrm{PTF}$ band. The model flux, adopting 
  best-fit parameters $\alpha=0.98$ and $t_0=-17.94\,\textrm{d}$, 
  is shown as a solid black line in the top panel. The lower panel 
  shows the model residuals, and the symbols are the same as in
  Figure~\ref{fig:lightcurve}. The model is fit to just the 
  $g_\mathrm{PTF}$-band observations to avoid systematic difference 
  between the filters (see the text for further 
  details). The joint distribution of $t_0$ and $\alpha$ is 
  illustrated in the inset of the upper panel. The solid and 
  dashed contours represent the $68\%$ and $99.7\%$ confidence 
  levels. \amiller{Update figure to remove other photometry}
  }
  \label{fig:early_lc_fit}
\end{figure}
